---
title: "Assignment 3 - Analysis of Power"
author: "Daija Odom"
date: "2023-03-01"
output: html_document
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Attach packages
library(tidyverse)
library(purrr)
library(ggpubr)
```

For the assignment you will write a new RMarkdown document based on analysis of power required for different possible_cars from looping.RMD. You will:

* Add an additional super light car with mass 5,000 kg and a surface area 10 m^2^

* Compute the mean and maximum power for Cars A, B, C and for your new car, assuming that average highway speed is 80 km/hr with a standard deviation of 10km/hr; Use 2 different methods to compute the mean and maximum power for each car type; a) FOR loop and b) the pmap function from purr

* Create two box plots (one for each method (FOR and pmap)) that show the range of power consumption (across sampled speeds) for each car type.


# Our highway

4 car types 

Imagine we have 4 different car types - and we know how often each occurs:

* car A  mass 20,000 kg, area 22 m^2^

* car B mass 65,000 kg, area 30 m^2^

* car C mass 38,000 kg area 22 m^2^

* car D mass 5,000 kg and a surface area 10 m^2^

Mean highway speed is 80 km/hr with a standard deviation of 10 km/hr

```{r}
# Set your working directory wherever your R functions are by clicking the "more" button in the right hand bottom corner OR use the here package
source("../R/autopower.R")

```

```{r}
# Add an additional super light car with mass 5,000 kg and a surface area 10 m^2^
# Generate a structure to store info on our possible cars
possible_cars = data.frame(name = c("A","B","C", "D"),mass=c(10000,65000,38000, 5000), area = c(22,30,22,10))

# Look at how results vary for mean speed say 80km/hr
# Do conversion from km/hr to m/s
speed_base = 80 * 0.278

# because I have one mass and area for each car and only 1 speed
# I can estimate power for each car type
# add to the data structure
possible_cars$power = autopower(V=speed_base, A = possible_cars$area, m=possible_cars$mass)

head(possible_cars)

# show results                         
ggplot(possible_cars, aes(x=mass, y=power, fill=as.factor(area)))+
  geom_col()+labs(y="Power (W)", x="Mass (kg)", fill="Area (m2)")

```

```{r}
# Use 2 different methods to compute the mean and maximum power for each car type; 
## a) FOR loop

# I want to estimate average power use given each car

possible_cars

# use sample to generate a distribution of speeds

# assume a log normal distribution of speeds with mean 80km/hr
# recall our function needs speed in m/s not km/hr so we will also do a conversion
# lets get a sample of a 200 speeds

nsample = 200
mean_speed = log(80*0.278)

speeds = rlnorm(mean=mean_speed, sd=0.125*mean_speed, nsample)
summary(speeds)

plot(density(speeds), ylab="Distribution of Speeds in (m/s)")

# we use a for loop to cycle through
# we need to create a data frame to store results - as above
# one column for each car and one row for each speed

powerall_1 = as.data.frame(matrix(nrow=length(speeds), ncol=nrow(possible_cars)))
# because we don't initialize it - values are NA
head(powerall_1)

# how many cars area there
nrow(possible_cars)
# for loops use an index - in this case "i" but you could use anything - it repeats
# anything between the {} for each values of i between 1 and nrow(possible_car) (which is 4 in our case)

# index in to a matrix (like powerall_1) is by row and column powerall_1[2,5] is 2nd row and 5th column
for (i in 1:nrow(possible_cars)) {
  powerall_1[,i] = autopower(A=possible_cars$area[i], m=possible_cars$mass[i], V=speeds)
}

# now it looks like above
head(powerall_1)
# we can add column names
colnames(powerall_1)=possible_cars$name

# plotting is the same as above

# for plotting sometimes its useful to turn columns in to rows
powerall_r1 = powerall_1 %>% pivot_longer(cols=everything(), names_to="car", values_to="power")
head(powerall_r1)

# quick visualization
method1_plot = ggplot(powerall_r1, aes(car,power, fill=car))+
  geom_boxplot()+
  ggtitle("Range of Power Consumption - For Loop")

```

```{r}
## b) The pmap function from purr
# the first part, generating speeds is the same
# what is I want to estimate average power use given  each car

possible_cars

# the first part is the same as above
# use sample to generate a distribution of speeds

# assume a log normal distribution of speeds with mean 100km/hr
# recall our function needs speed in m/s not km/hr so we will also do a conversion
# lets get a sample of a 200 speeds- we could also do this by actually measuring speeds

nsample = 200
mean_speed = log(80*0.278)

speeds = rlnorm(mean=mean_speed, sd=0.125*mean_speed, nsample)
summary(speeds)

plot(density(speeds), ylab="Distribution of Speeds in (m/s)")

# how do we run each car for all speeds 
# pmap runs a function for each value in a list of parameters, with other parameters set for each iteration

powerall_2 = pmap(list(A = possible_cars$area, m=possible_cars$mass), autopower, V=speeds)

# lets turn to a data frame for easier graphing
# we can add column names
powerall_2 = as.data.frame(powerall_2, col.names=possible_cars$name)

head(powerall_2)

# for plotting sometimes its useful to turn columns in to rows
powerall_r2 = powerall_2 %>% pivot_longer(cols=everything(), names_to="car", values_to="power")
head(powerall_r2)

# quick visualization

method2_plot = ggplot(powerall_r2, aes(car,power, fill=car))+
  geom_boxplot()+
  ggtitle("Range of Power Consumption - pmap")
method2_plot

ggarrange(method1_plot, method2_plot, nrow=2)

powerall_1 %>% map(mean)
powerall_2 %>% map(mean)

```

